---
title: "SC_data"
author: "Riddhi Mehta-Neugebauer"
date: "8/20/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Load Libraries
```{r}
# let's call out specific packages we need rather than just loading tidyverse (I find it helps people to know what packages they're using)
library(readr) # for reading csvs
library(dplyr) # for data management
library(stringr) # for working with strings
library(purrr) # for working with lists
```


# Load Data
```{r}
# list the files in the folder
files <- list.files(path = "./CRDC_files", pattern = "*.csv", full.names = TRUE)


# apply read_csv to all the files listed in files
data <- lapply(files, read_csv) # using readr
```

# Convert List to Dataframe
## Rename dataframes in the list

```{r}
# rename the list elements to represent the filename
data <- setNames(data, stringr::str_replace(list.files(path = "./CRDC_files", pattern = "*.csv"), "\\.csv", "")) # running same function again but without the breadcrumb trail

# check the list to see the names of the dataframes
names(data)
```

## Extract only STEM data frames
```{r}
# list the STEM filenames
key_file_names <- c("Advanced Mathematics", "Algebra II",  "Biology", "Calculus", "Chemistry", "Geometry", "Physics")

data_stem <- data[key_file_names]

```

## Extract the male and female enrollment columns for each of the STEM fields and combine into one dataframe
```{r}
# what are the column names in Calculus for example
 names(data_stem[["Calculus"]])

# Check with the codebook to see what code for the variable you are interested in is. I want total enrollment in each STEM class for males and females. Note the column names you are most interested in. 

key_col_names <- c("LEA_STATE", "LEA_STATE_NAME", "LEAID", "LEA_NAME", "SCHID", "SCH_NAME", "COMBOKEY", "JJ", "TOT_SCIENR_BIOL_M", "TOT_SCIENR_BIOL_F", "TOT_MATHENR_CALC_M", "TOT_MATHENR_CALC_F", "TOT_SCIENR_CHEM_M", "TOT_SCIENR_CHEM_F", "TOT_MATHENR_ALG2_M", "TOT_MATHENR_ALG2_F", "TOT_MATHENR_GEOM_M", "TOT_MATHENR_GEOM_F", "TOT_SCIENR_PHYS_M", "TOT_SCIENR_PHYS_F", "TOT_MATHENR_ADVM_M", "TOT_MATHENR_ADVM_F")

# Apply the select function on all the sublists
data_stem <- lapply(data_stem, function(x) x %>% select(any_of(key_col_names)))


```

## Merge the lists into a single dataframe, based on COMBOKEY (LEAID + SCHID)
```{r}
data_merge <- data_stem %>% 
  reduce(inner_join, by = c("LEA_STATE", "LEA_STATE_NAME", "LEAID", "LEA_NAME", "SCHID", "SCH_NAME", "JJ", "COMBOKEY"))
```

## Replace missing values with NA
```{r}
data_merge <- na_if(data_merge, "-9") # -9 = Not Applicable / Skipped
data_merge <- na_if(data_merge, "-6") # -6 = Force Certified
data_merge <- na_if(data_merge, "-5") # -5 = Action Plan

summary(data_merge)
```

## Remove obs. with all NA values (indicator of non-high school)
```{r}
# Total the number of enrollments in STEM fields
data_merge <- data_merge %>% 
  mutate(enroll_sum = select(., TOT_MATHENR_ADVM_M, TOT_MATHENR_ADVM_F, TOT_MATHENR_ALG2_M, TOT_MATHENR_ALG2_F, TOT_SCIENR_BIOL_M, TOT_SCIENR_BIOL_F,  TOT_MATHENR_CALC_M, TOT_MATHENR_CALC_F, TOT_SCIENR_CHEM_M,  TOT_SCIENR_CHEM_F,  TOT_MATHENR_GEOM_M, TOT_MATHENR_GEOM_F, TOT_SCIENR_PHYS_M,  TOT_SCIENR_PHYS_F) %>% rowSums(na.rm = T))

# Filter out the enroll_sum = 0 and those that aren't juvenile justice facility
data_merge2 <- data_merge %>% 
  filter(enroll_sum != 0 & JJ == "No")

```


# Analyze Data
```{r}
# 
```

